---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
summary <- read_rds("summarySCC_PM25.rds") %>%as_tibble(rownames = "tons") %>% mutate(tons=as.numeric(tons))
summary
```

```{r}
source_classification <- read_rds("Source_Classification_Code.rds") %>% as_tibble()
source_classification
```
```{r}
summary$year %>% unique()
```
#Have total emissions from PM2.5 decreased in the United States from 1999 to 2008? Using the base plotting system, make a plot showing the total PM2.5 emission from all sources for each of the years 1999, 2002, 2005, and 2008.
```{r}
p1 <- summary %>% group_by(year) %>% summarise(tons_total=sum(tons)) %>% ggplot(aes(year,tons_total))+geom_col()


ggsave("plot1.png", plot = p1)
p1
```
## Theres an increase over the years

#Have total emissions from PM2.5 decreased in the Baltimore City, Maryland (\color{red}{\verb|fips == "24510"|}fips == "24510") from 1999 to 2008? Use the base plotting system to make a plot answering this question.
```{r}
p2 <- summary %>% filter(fips=="24510") %>% group_by(year) %>% summarise(tons_total=sum(tons)) %>% ggplot(aes(year,tons_total))+geom_col() 
p2
```
##yes emmisions have also increased in Baltimore
#
#

#Of the four types of sources indicated by the (point, nonpoint, onroad, nonroad) variable, which of these four sources have seen decreases in emissions from 1999–2008 for Baltimore City? Which have seen increases in emissions from 1999–2008? Use the ggplot2 plotting system to make a plot answer this question.
```{r}
p3 <- summary%>% filter(fips=="24510") %>% group_by(year, type) %>% summarise(tons_total=sum(tons)) %>% ggplot(aes(year,tons_total, color=type))+geom_line()
p3
```

## Answer 2) all have incresed
#
#
#Across the United States, how have emissions from coal combustion-related sources changed from 1999–2008?
```{r}
comb <- source_classification[str_detect(source_classification$Short.Name, pattern="Comb"),]

p4 <- left_join(comb,summary)%>% group_by(year) %>% summarise(tons_total=sum(tons)) %>% ggplot(aes(year,tons_total))+geom_line() + ggtitle("emissions from coal combustion-related sources")

p4

```
##emissions from coal combustion-related sources have risen


#
#
#How have emissions from motor vehicle sources changed from 1999–2008 in Baltimore City?
```{r}

p5 <- source_classification %>% left_join(summary)%>% 
  group_by(year) %>% 
  filter(EI.Sector %in% str_subset(EI.Sector,fixed('vehicle', ignore_case=TRUE))) %>% 
  filter(fips=="24510") %>% 
  summarise(tons_total=sum(tons)) %>% ggplot(aes(year,tons_total))+geom_line() + ggtitle("emissions from motor vehicle sources changed from 1999–2008 in Baltimore City")
p5

```
##they have quita dramtically risen

#Compare emissions from motor vehicle sources in Baltimore City with emissions from motor vehicle sources in Los Angeles County, California (\color{red}{\verb|fips == "06037"|}fips == "06037"). Which city has seen greater changes over time in motor vehicle emissions?
```{r}
p6 <- source_classification %>% left_join(summary)%>% 
  filter(EI.Sector %in% str_subset(EI.Sector,fixed('vehicle', ignore_case=TRUE))) %>% 
  filter(fips%in% c("24510","06037")) %>% mutate(city=ifelse(fips=="24510","Baltimore","L.A.")) %>%  group_by(year,city) %>% 
  summarise(tons_total=sum(tons)) %>% ggplot(aes(year,tons_total, color=city))+geom_line() + ggtitle("emissions from motor vehicle sources changed from 1999–2008 in Baltimore City and LA")
p6
```

#LA could reduce its emissions in the last years